---
title: "The Evolution of AI and Robotics"
subtitle: "How trial-and-error learning turns robot-arm actions into skills"
author: ""
format:
  html:
    toc: true
    theme: cosmo
    code-fold: true
execute:
  warning: false
  message: false
---

```{r}
library(ggplot2)
library(dplyr)
library(scales)

set.seed(2026)

# This is a reproducible illustration, not a download of the benchmark data.
task_info <- tibble::tibble(
  task = c("Push", "Pick and place", "Slide", "Grasp"),
  difficulty = c(0.63, 0.48, 0.56, 0.38),
  color = c("#E76F51", "#2A9D8F", "#F4A261", "#264653")
)

evolution_stages <- tibble::tibble(
  stage = 1:4,
  title = c("Human instructions", "Simulation", "AI policy", "Robot skill"),
  detail = c("Rules written by people", "Safe practice world", "Learns from outcomes", "Acts in the world"),
  color = c("#264653", "#2A9D8F", "#E9C46A", "#E76F51")
)

episodes <- tidyr::crossing(
  task = task_info$task,
  attempt = 1:120
) |>
  left_join(task_info, by = "task") |>
  mutate(
    learning = 0.18 * log1p(attempt) / log(121),
    probability = pmin(0.96, difficulty + learning),
    success = rbinom(n(), size = 1, prob = probability)
  )

task_summary <- episodes |>
  group_by(task, color) |>
  summarise(
    successes = sum(success),
    trials = n(),
    success_rate = successes / trials,
    standard_error = sqrt(success_rate * (1 - success_rate) / trials),
    .groups = "drop"
  ) |>
  mutate(
    lower = pmax(0, success_rate - 1.96 * standard_error),
    upper = pmin(1, success_rate + 1.96 * standard_error)
  )

learning_curve <- episodes |>
  mutate(window = ((attempt - 1) %/% 10) * 10 + 10) |>
  group_by(task, color, window) |>
  summarise(success_rate = mean(success), .groups = "drop")

plot_theme <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 17, color = "#17324D"),
    plot.subtitle = element_text(color = "#496579"),
    axis.title = element_text(face = "bold", color = "#17324D"),
    axis.text = element_text(color = "#496579"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(12, 18, 12, 12)
  )
```

## From programmed motion to learned skill

Robotics data often arrives as a **simulation environment** rather than a rectangular table. Each episode can contain observations, actions, rewards, and a final success signal. That makes the first useful question: *what does one episode look like, and how should success be summarized?*

The OpenAI Fetch and Hand environments are a promising setting for that question. They include robot-arm behaviors such as pushing and grasping, which makes them useful for studying a major shift in robotics: moving from instructions written entirely by people toward AI policies that improve through repeated attempts.

The figures below are a **synthetic demonstration** because the repository does not yet contain episode-level Fetch or Hand data. The structure mirrors the kind of summary we could make after loading real simulation episodes.

## The evolution of an intelligent robot

```{r}
evolution_links <- tibble::tibble(
  x = 1:3,
  xend = 2:4,
  y = 0.5,
  yend = 0.5
)

ggplot() +
  geom_segment(
    data = evolution_links,
    aes(x = x, xend = xend, y = y, yend = yend),
    arrow = grid::arrow(length = grid::unit(0.18, "inches"), type = "closed"),
    color = "#9BB3C4",
    linewidth = 1.2
  ) +
  geom_point(
    data = evolution_stages,
    aes(x = stage, y = 0.5, fill = color),
    shape = 21,
    size = 14,
    color = "white",
    stroke = 2.5
  ) +
  geom_text(
    data = evolution_stages,
    aes(x = stage, y = 0.82, label = title, color = color),
    fontface = "bold",
    size = 4.1
  ) +
  geom_text(
    data = evolution_stages,
    aes(x = stage, y = 0.16, label = detail),
    color = "#496579",
    size = 3.5
  ) +
  scale_fill_identity() +
  scale_color_identity() +
  scale_x_continuous(breaks = evolution_stages$stage, labels = NULL, limits = c(0.65, 4.35)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  labs(
    title = "From instructions to intelligence",
    subtitle = "AI gives robots a way to improve inside a simulated world before acting in the real one",
    x = NULL,
    y = NULL
  ) +
  plot_theme +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(24, 18, 24, 18)
  )
```

## Which tasks are succeeding?

```{r}
ggplot(task_summary, aes(x = reorder(task, success_rate), y = success_rate, fill = color)) +
  geom_col(width = 0.66, show.legend = FALSE) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.10, color = "#17324D", linewidth = 0.7) +
  geom_text(aes(label = percent(success_rate, accuracy = 1)), hjust = -0.12,
            color = "#17324D", fontface = "bold", size = 4.2) +
  coord_flip(clip = "off") +
  scale_fill_identity() +
  scale_y_continuous(labels = percent, limits = c(0, 1.08), breaks = seq(0, 1, 0.2)) +
  labs(
    title = "Different robot skills evolve at different speeds",
    subtitle = "Illustrative success rates after 120 simulated AI-policy episodes; bars show 95% intervals",
    x = NULL,
    y = "Successful episodes"
  ) +
  plot_theme
```

This view is useful for comparing behaviors, but it hides *when* the successes happened. A robot that succeeds only after many failed attempts tells a different story from one that succeeds immediately.

## The evolution happens through repetition

```{r}
ggplot(learning_curve, aes(x = window, y = success_rate, color = color, group = task)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "#A8B7C3") +
  geom_line(linewidth = 1.25) +
  geom_point(size = 2.7) +
  facet_wrap(~ task, nrow = 1) +
  scale_color_identity() +
  scale_x_continuous(breaks = c(10, 60, 120), expand = expansion(mult = c(0.04, 0.08))) +
  scale_y_continuous(labels = percent, limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  labs(
    title = "AI turns trial and error into a robotics learning curve",
    subtitle = "Each point represents a later stage in the policy's evolution",
    x = "Episodes attempted",
    y = "Success rate"
  ) +
  plot_theme +
  theme(strip.text = element_text(face = "bold", color = "#17324D", size = 11))
```

The upward pattern is the visual story of AI and robotics evolution: experience can turn uncertain actions into more reliable behavior. With the actual dataset, the next step would be to replace `episodes` with a table created from the environment logs, then compare success by task, seed, and training method.

## What this makes me wonder

It is remarkable that an AI policy can turn trial and error into physical skill. This is the larger evolution: robotics supplies the body, AI supplies an adaptable decision-maker, and simulation supplies a place to learn safely. But the same progress raises a human question: as robots take on more tasks, how will our work change, and what new kinds of work will we choose to do? A good final project can hold both ideas at once: measure what the robot learns, and ask what that learning means for people.

::: {.callout-note}
### Data note

The plotted values are simulated with a fixed seed for demonstration. They should not be reported as findings about the OpenAI benchmark. For a final analysis, document the exact environment version, task definition, number of episodes, random seeds, and success metric.
:::